Automated Example Oriented REST API Documentation

### 2017-11-08 introduction.tex
final version

Robillard defines an API as follows: An API is the interface to implement functionality that developers can access to perform various tasks \cite{Robillard_a_field_study, Robillard_what_makes}. Application Programming Interfaces, commonly known as APIs, are used to express a software component in terms of its operations, inputs, outputs, and their types\footnote{\url{https://en.wikipedia.org/wiki/Application_programming_interface}}. APIs enable multiple software components to interact with each other. Web APIs allow software-to-software interconnectivity using web technologies as the transport mechanism between systems.

REST APIs are a type of API that are used to integrate software components using web technologies. Fielding defined Representational State Transfer or REST as an architectural style for developing distributed hypermedia systems \cite{Fielding_rest}. He described the design of modern Web as an example of a REST-style architecture. REST-style architecture allows software components to satisfy the data transfer needs of the modern Internet-scale applications by leveraging a generic interface based on URIs and Internet media types that allow intemediate processing such as caching to improve system performance.

REST APIs allow access to application resources over HTTP. While REST describes a generic protocol for accessing application resources, the applications are responsible for the specific representation of the resources. For example, a REST API may support multiple Internet media types such as JSON, XML, CSV for a single application resource. Similarly, two different REST APIs may use the same media type but contain entirely different formats to represent their objects.

@@

Previous research in the area of API usability mostly focused on local APIs such as Java libraries. Researchers identified the documentation of APIs as both the primary source of information as well as the key obstacle for API usability \cite{Robillard_what_makes}. Hence the quality of the API documentation plays an important role in API usability. To this regard, researchers have identified the qualities of ``good API documentation'' as follows: complete, correct, includes thorough explanations and code examples, provides consistent presentation and organization \cite{Robillard_what_makes,Myers_study}. Today, there are several tools such as Javadoc\footnote{\url{http://www.oracle.com/technetwork/java/javase/documentation/index-jsp-135444.html}}, UsETeC \cite{zhu2014mining}, Jadeite \cite{Jadeite}, APIMiner \cite{montandon2013documenting}, Roast \cite{Hoffman_api_documentation} that can be used to document local APIs with the aforementioned qualities .

@@

Currently, the process of documenting REST APIs is largely manual. Conceptually, the high level steps involved in the manual process can be identified as shown in Figure \ref{fig:manual}. These steps 1-7 can be described as follows:

@@

\section{Research Goals}
The overarching goal of my research is to improve the current state of the process of REST API documentation process by automating the repeating manual steps in the aforementioned workflow. More specifically, I aim to solve the following research goals:

\begin{itemize}
  \item \textit{RG1: To identify a set of REST API documentation requirements.} Existing research on API usability has derived a number of API documentation requirements primarily focusing on local APIs. Due to the structural and implementation differences between local and REST APIs, I aim to identity REST API specific documentation requirements in my research. This has several implications as follows: having access to a scientifically derived list of REST API documentation requirements allows REST API developers to make an informed decision related to planning and execution of their documentation efforts. The requirements provide a background research that can be used by practitioners to compare and contrast existing REST API documentation tools while selecting an appropriate tool for their API. REST API documentation tool developers can also use the requirements to build necessary features.
  \item \textit{RG2: To develop and evaluate a technique to overcome the commonly observed challenges related to REST API documentation.} I aim to devise a reusable technique so that REST APIs can be documented following the identified REST API documentation requirements without following manual or bespoke methods. It has the following implications: REST API developers can utilize the technique to overcome the common challenges with the generation and maintenance of their API documentation, and REST API client developers can access an API documentation.
\end{itemize}

\section{Research Objectives}
Related to RG1 and RG2, I aim to address the following research questions as my objectives:

\begin{itemize}
  \item \textit{RO1: How are evolving REST APIs documented?} This question is aimed to analyze the gap between current industry practice around the documentation of evolving REST APIs and the propose solutions from existing literature. This objective helped me to identify the research opportunity to work on unresolved problems related to REST API documentation.
  \item \textit{RO2: What are the requirements for usable REST API documentation?} This objective provides a necessary background for improving the state of REST API documentation process.
  \item \textit{RO3: What technique can be used to document REST APIs to meet the requirements?} This is the central research objective. I researched to develop a new technique that can fill the gap between existing REST API documentation methods and the identified requirements.
  \item \textit{RO4: Is the proposed technique a practical one for REST API developers?} A qualitative evaluation of the proposed technique in a real-world setup is needed to understand the feasibility and the shortcomings of the proposed technique from the perspective of the REST API developers.
  \item \textit{RO5: How does a documentation using the proposed technique impact REST API client developers?} This question investigates the impact of the proposed technique from the perspective of the REST API client developer effort and success rate while performing API tasks.
\end{itemize}

\section{Cohesion and Unitary Report}
To address the aforementioned research goals and objectives, my research strategy included five sub-projects, have resulted in publications at peer-reviewed international conferences. Table~\ref{table:chapter_summary} shows a high level overview of the flow of arguments between the sub-projects to meet the research goals.

\begin{table}
  \caption{Summary and Flow of Arguments of the Sub-projects}
  \begin{tabular}{|p{1.5cm}|p{1.5cm}|p{1.5cm}|p{4.5cm}|p{6cm}|}
  \hline
  Chapter & Goal & Objective & Research Project & Primary Outcome \\
  \hline
  Chapter~\ref{chapter:case_study}  & RG1 & RO1 & A case study of Web API evolution \cite{sohan2015case}. & A set of challenging problems related to evolving Web APIs. Selected REST API documentation as the problem of interest for this research. \\
  \hline
  Chapter~\ref{chapter:spy_rest} & RG1 and RG2 & RO2 and RO3 & SpyREST: Automated RESTful API documentation using an HTTP proxy server \cite{sohan2015spyrest}. & A list of requirements. A conceptual design using interception to generate REST API documentation to meet the requirements. \\
  \hline
  Chapter~\ref{chapter:demo_paper} & RG2 & RO3 & SpyREST in action: An automated RESTful API documentation tool \cite{sohan2015spyrest_tool}. & A reusable tool that implements the conceptual design and shows how the identified REST API documentation requirements are met.\\
  \hline
  Chapter~\ref{chapter:cisco_study} & RG2 & RO4 & Automated example oriented REST API documentation at Cisco \cite{sohan_cisco}. & Empirical evidence that the proposed technique provides a practical documentation solution for REST API developers.\\
  \hline
  Chapter~\ref{chapter:controlled_study} & RG2 & RO5 & A study of the effectiveness of usage examples in REST API documentation \cite{sohan_vlhcc}. & Empirical evidence that the proposed technique allows REST API client developers to perform API tasks with less effort, higher success and satisfaction ratings.\\
  \hline
\end{tabular}
\label{table:chapter_summary}
\end{table}

@@

I discuss my fourth research project, related to RG2, involving an industrial evaluation of the proposed technique and the tool in Chapter~\ref{chapter:cisco_study} \cite{sohan_cisco}. The evaluation is carried out based on the data collected from an eighteen month of production use of SpyREST at Cisco, my current employer. From the study I found that it is feasible to use the proposed interception technique and automated tests to document evolving REST APIs satisfying the previously presented REST API documentation requirements. Automated always-updated REST API documentation was found to help establish a quick feedback cycle among the stakeholders. I also discuss several refinements and limitations of the proposed technique based on the lessons learned from this case study.

@@

In Chapter~\ref{chapter:conclusion}, I conclude by discussing how the research objectives are achieved and provide a list of problems for future research.

@@

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,references}

>>>>

Application Programming Interfaces, commonly known as APIs, are used to express a software component in terms of its operations, inputs, outputs, and their types\footnote{\url{https://en.wikipedia.org/wiki/Application_programming_interface}}. Robillard defines an API as follows: An API is the interface to implement functionality that developers can access to perform various tasks \cite{Robillard_a_field_study, Robillard_what_makes}. APIs enable multiple software components to interact with each other. Web APIs allow software-to-software interconnectivity using web technologies as the transport mechanism between systems.

REST APIs are a type of API that are used to integrate software components using web technologies. Fielding defined Representational State Transfer or REST as an architectural style for developing distributed hypermedia systems \cite{Fielding_rest}. Fielding described the design of the modern Web as an example of a REST-style architecture. REST-style architecture allows software components to satisfy the data transfer needs of the modern Internet-scale applications by leveraging a generic interface based on URIs and Internet media types that allow intemediate processing such as caching to improve system performance.

The overarching goal of my research is to improve the current state of the REST API documentation process by automating the repeating manual steps. REST APIs allow access to application resources over HTTP. While REST describes a generic protocol for accessing application resources, the applications are responsible for the specific representation of the resources. For example, a REST API may support multiple Internet media types such as JSON, XML, CSV for a single application resource. Similarly, two different REST APIs may use the same media type but contain entirely different formats to represent their objects.

@@

Previous research in the area of API usability mostly focused on local APIs such as Java libraries. Researchers identified the documentation of APIs as both the primary source of information as well as the key obstacle for API usability \cite{Robillard_what_makes}. Hence the quality of the API documentation plays an important role in API usability. To this regard, researchers have identified the qualities of ``good API documentation'' as follows: complete, correct, includes thorough explanations and code examples, provides consistent presentation and organization \cite{Robillard_what_makes,Myers_study}. Today, there are several tools such as Javadoc\footnote{\url{http://www.oracle.com/technetwork/java/javase/documentation/index-jsp-135444.html}}, UsETeC \cite{zhu2014mining}, Jadeite \cite{Jadeite}, APIMiner \cite{montandon2013documenting}, Roast \cite{Hoffman_api_documentation} that can be used to document local APIs with the aforementioned qualities.

@@

Currently, the process of documenting REST APIs is largely manual. This is problematic because, it is time a consuming and an error-prone process. Conceptually, the high level steps involved in the manual process can be identified as shown in Figure \ref{fig:manual}. These steps 1-7 can be described as follows:

@@

\section{Research Goals}
The principal goal of my research is to improve the REST API documentation process by automating the repeating manual steps in the aforementioned workflow. More specifically, I aim to solve the following research goals:

\begin{itemize}
  \item \textit{RG1: To identify a set of REST API documentation requirements.} Existing research on API usability has derived a number of API documentation requirements primarily focusing on local APIs. Due to the structural and implementation differences between local and REST APIs, I aim to identity REST API specific documentation requirements in my research. This has several implications as follows: having access to a scientifically derived list of REST API documentation requirements allows REST API developers to make an informed decision related to planning and execution of their documentation efforts. The requirements provide a background research that can be used by practitioners to compare and contrast existing REST API documentation tools while selecting an appropriate tool for their API. REST API documentation tool developers can also use the requirements to build necessary features.
  \item \textit{RG2: To develop and evaluate a documentation technique to overcome the commonly observed challenges related to REST API documentation.} I aim to devise a reusable technique so that REST APIs can be documented following the identified REST API documentation requirements without following manual or bespoke methods. It has the following implications: REST API developers can utilize the technique to overcome the common challenges with the generation and maintenance of their API documentation, and REST API client developers can access an API documentation.

\end{itemize}

\section{Research Questions}
Related to RG1 and RG2, I aim to address the following research questions as my objectives:

\begin{itemize}
  \item \textit{RQ1: How are evolving REST APIs documented?} This question is aimed to analyze the gap between current industry practice around the documentation of evolving REST APIs and the propose solutions from existing literature. This objective helped me to identify the research opportunity to work on unresolved problems related to REST API documentation.
  \item \textit{RQ2: What are the requirements for usable REST API documentation?} Answer of this question provides a necessary background for improving the state of REST API documentation process.
  \item \textit{RQ3: What technique can be used to document REST APIs to meet the requirements?} This is the central research question. I researched to develop a new technique that can fill the gap between existing REST API documentation methods and the identified requirements.
  \item \textit{RQ4: Is the proposed technique practical for REST API developers?} A qualitative evaluation of the proposed technique in a real-world setup is needed to understand the feasibility and the shortcomings of the proposed technique from the perspective of the REST API developers.
  \item \textit{RQ5: How does documentation using the proposed technique impact REST API client developers?} This question investigates the impact of the proposed technique from the perspective of the REST API client developer effort and success rate while performing API tasks.
\end{itemize}

\section{Cohesion and Unitary Report}
To address the aforementioned research goals and objectives, my research strategy included five sub-projects, which have resulted in publications at peer-reviewed international conferences. Table~\ref{table:chapter_summary} shows a high level overview of the flow of arguments between the sub-projects to meet the research goals and questions.

\begin{table}
  \caption{Summary and Flow of Arguments of the Sub-projects}
  \begin{tabular}{|p{1.5cm}|p{1.5cm}|p{1.5cm}|p{4.5cm}|p{6cm}|}
  \hline
  Chapter & Goal & Question& Research Project & Primary Outcome \\
  \hline
  Chapter~\ref{chapter:case_study}  & RG1 & RQ1 & A case study of Web API evolution \cite{sohan2015case}. & A set of challenging problems related to evolving Web APIs. Selected REST API documentation as the problem of interest for this research. \\
  \hline
  Chapter~\ref{chapter:spy_rest} & RG1 and RG2 & RQ2 and RQ3 & SpyREST: Automated RESTful API documentation using an HTTP proxy server \cite{sohan2015spyrest}. & A list of requirements. A conceptual design using interception to generate REST API documentation to meet the requirements. \\
  \hline
  Chapter~\ref{chapter:demo_paper} & RG2 & RQ3 & SpyREST in action: An automated RESTful API documentation tool \cite{sohan2015spyrest_tool}. & A reusable tool that implements the conceptual design and shows how the identified REST API documentation requirements are met.\\
  \hline
  Chapter~\ref{chapter:cisco_study} & RG2 & RQ4 & Automated example oriented REST API documentation at Cisco \cite{sohan_cisco}. & Empirical evidence that the proposed technique provides a practical documentation solution for REST API developers.\\
  \hline
  Chapter~\ref{chapter:controlled_study} & RG2 & RQ5 & A study of the effectiveness of usage examples in REST API documentation \cite{sohan_vlhcc}. & Empirical evidence that the proposed technique allows REST API client developers to perform API tasks with less effort, higher success and satisfaction ratings.\\
  \hline
\end{tabular}
\label{table:chapter_summary}
\end{table}

@@

I discuss my fourth research project, related to RG2, involving an industrial evaluation of the proposed technique and the tool in Chapter~\ref{chapter:cisco_study} \cite{sohan_cisco}. The evaluation is carried out based on the data collected from an 18 month production use of SpyREST at Cisco, my current employer. From the study I found that it is effective to use the proposed interception technique and automated tests to document evolving REST APIs satisfying the previously presented REST API documentation requirements. Automated always-updated REST API documentation was found to help establish a quick feedback cycle among the stakeholders. I also discuss several refinements and limitations of the proposed technique based on the lessons learned from this case study.

@@

In Chapter~\ref{chapter:conclusion}, I conclude by discussing how the research objectives are achieved and provide a list of potential future research problems.

@@

\newpage
\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,references}

### 2017-11-08 root.tex
adds a line about client library
final version

\include{introduction}
\include{case_study/case_study}
\include{spy_rest/spy_rest}
\include{demo_paper/demo_paper}
\include{cisco_study/cisco_study}
\include{controlled_study/controlled_study}
\chapter{Conclusion}
\label{chapter:conclusion}

In this thesis, I presented a set of five primary research objectives related to the topic of REST API documentation. I have addressed the first research objective, RO1, of finding the gap between the current industry practices and existing research on REST API documentation by performing a case study in my first research sub-project. I started by reviewing the existing literature on Web API evolution. From the existing empirical research on Web API evolution, I found that researchers have focused on profiling SOAP based web services to identify the common change patterns. They suggested focusing on the maintenance efforts related to Web API evolution because of its lingering effects. In the existing literature, I found several techniques proposed by researchers related to supporting the evolution of SOAP APIs. Analyzing the existing literature on tool support for Web API evolution, I found tools developed by researchers that primarily rely on custom formatted API specification documents that require a manual effort to generate and maintain. In this sub-project, I have also analyzed the evolution of nine Web APIs in production, selected based on their industry, popularity, maturity, and source code access, to understand the current state of practice around Web API versioning, documentation, and change communication patterns. By comparing and contrasting the findings from the literature review and the industry practices, I observed that it was important to carry out further research to find techniques to support versioning, documentation, and change communication needs for evolving REST APIs.

I addressed RO2, finding REST API documentation requirements, in my second research sub-project. I started by analyzing the existing literature on API usability to summarize a list of general API documentation requirements as derived by other researchers. Then, I analyzed the existing literature on REST API documentation. Based on these two analyses, I derived the following list of documentation requirements specific for evolving REST APIs: automated, example based, executable, version aware, customizable, reusable, and collaborative. These requirements are used as a background work to address my subsequent research objectives.

RO3, the primary objective of finding a new technique to satisfy the documentation requirements of evolving REST APIs, is addressed in my third and fourth sub-projects. I observed that most existing techniques for REST API documentation relied on custom API specification formats as an input. This approach requires manual effort to generate and maintain the API specification as a separate artifact with the API as the API evolves and fails to satisfy the identified REST API documentation requirements. I proposed a novel technique called SpyREST to automatically intercept and transform example REST API calls into usable API documentation following the identified requirements. An HTTP proxy server is used to collect data from example API calls. The collected data is then synthesized to automatically infer the structural elements of the API such as the version, API resource, HTTP verb, API action, query parameters, request and response body, and HTTP headers. In addition to the structure, executable API examples are automatically re-constructed with realistic data from the recorded API calls. I developed a prototype tool to demonstrate a concrete implementation of the proposed concept. In a preliminary case study with three evolving REST APIs, I have shown that it is possible to use the proposed technique to generate a more accurate documentation of REST APIs compared to their official documentation that are  manually maintained.

I addressed RO4, an evaluation of the proposed technique from the perspective of REST API developers, in my fourth sub-project using an industrial case study. I presented the proposed technique and the prototype tool to a team of software engineers at Cisco when we were in the process of selecting a documentation method for a suite of new REST APIs to be built. I discussed the selection process behind choosing SpyREST as documentation tool for this project. In this case study, I discussed how the proposed technique of using test code to drive REST API documentation has been leveraged in Cisco to create and maintain an always-updated and accurate documentation for an evolving REST API in production for a period of eighteen months. I gathered feedback from the stakeholders including the API developers, QA engineers, project managers, and customer support engineers to evaluate the merits and shortcomings of the proposed technique based on their production use. The evaluation results show that it is feasible to use API test code to generate and maintain REST API documentation for an evolving API with the identified REST API documentation requirements. From this evaluation, I also found that different stakeholders are able to establish a fast and continuous feedback loop about the API and its documentation while developing the API features when the API documentation is auto updated from executable code. Additionally, from this evaluation, I found a list of limitations of the proposed technique related to the lack of support for using version control features for custom edited API documentation content and automatically generated changelogs showing the differences between API elements of different API versions. Future research can be carried out to address these limitations.

My last research objective, RO5, to evaluate the impact of the proposed technique on REST API client developers, was addressed in my fifth sub-project. I performed a controlled study with 26 professional software engineers with prior REST API experience. The participants were asked to perform tasks using a REST API following its accompanying documentation. I enhanced an existing API documentation with usage examples leveraging the API's test code. This enhancement mimics the outcome of the proposed technique but does not use the proposed tool. Participants were mapped into two groups. One of the groups used the original API documentation and the other group used the enhanced documentation. I provided a web based API explorer so that participants could only focus on the required API inputs instead of writing a full API client application to perform the tasks. I observed that participants using the original API documentation faced similar patterns of obstacles involving the use of correct data types, data formats, HTTP headers, and request body. Participants using the enhanced documentation were able to reduce these obstacles to perform API tasks with a higher success rate, requiring less trial API calls and effort, and a higher satisfaction rating. This provides empirical evidence that the proposed technique has a positive impact on the REST API client developers. I also derived a list of recommendations related to the content of usage examples for REST API documentation that can be used by practitioners as a guideline to prioritize their documentation efforts. From this study, I also observed that participants used more trial API calls for API tasks that involve multiple dependent API calls even while using the enhanced documentation. Future research is needed to generate API documentation with examples for complex API usage scenarios to reduce this obstacle.

@@

\section{Future Work}
Based on my findings from the this research, I identified the following list of potential future research problems.

\begin{itemize}
  \item \textbf{Extension to Support Complex Usage Scenarios.} Future research needs to be carried out to find techniques to automate the process of documenting usage scenarios that require multiple API calls to perform complex API tasks. Future research is also needed to investigate techniques for automatically cross-referencing API elements to allow API client developers with an improved navigation experience while reading the REST API documentation. In Chapter~\ref{chapter:controlled_study}, I discussed that the study participants faced obstacles with performing API tasks that required the use of multiple related API calls to reach a desired outcome. Future research needs to investigate techniques that can be used to provide links to related API elements and usage examples to allow API client developers with an alternative navigation experience.
  \item \textbf{Automated Changelog Generation.} For frequently evolving APIs, future research needs to extend the proposed technique in this thesis to automatically compute human readable changelogs and detect version compatibility to follow a semantic versioning scheme. This would help API client developers with an accurate information about the potential impact of an API update on their own application. In Chapter~\ref{chapter:case_study}, I discussed that REST APIs evolve frequently, often several times per month, and the change profiles contain different types of changes. I also provided a list of change patterns that I observed while studying the change profiles. In Chapter~\ref{chapter:cisco_study}, I discussed that manual effort was required to maintain the API changelog for the studied API. Future research needs to investigate automated changelog generation to help the evolution of REST APIs. The outcome of an automated changelog generation can be applied to automatically infer a semantic version of a REST API based on the inferred compatibility information.
  \item \textbf{Automated API Client Code Generation.} Several tools exist that can convert custom REST API description files to API client code in multiple programming languages. Future research can be carried out to map the SpyREST data model with the custom API description languages to leverage the interception technique to reduce the amount of manual effort required to maintain the custom formatted API descriptions as the API evolves.
  \item \textbf{Extensions of the Interception Based Documentation to non-REST APIs.} The proposed technique can be extended to document usage examples for other stateless APIs outside the realm of REST APIs. The structure of REST APIs are commonly documented using programming language features and formatted comments. Using formatted coments to provide API documentation with executable code examples requires manual effort since no automated feedback is available about the accuracy of the comments. API developers commonly write and maintain API test code that shows the intended way of using the APIs. The API test code provides automated feedback by means of test results. However, API test code may not be a substitute for API documentation. API test code may contain complex setup, tear-down, and verification logic that are not useful for documentation. Also, for closed source applications, organizations may not be able to publish their source code to external API client developers. The concept of intereception as described in Chapter~\ref{chapter:spy_rest} can be applied to other forms of API to generate and maintain API documenation by intercepting REST API tests. For example, several programming languages support concepts such as Aspect Oriented Programming, where custom code can be written to intercept API method calls, that can be used for API documentation. Future research can investigate the feasibility of leveraging the interception technique to source realistic usage examples for local stateless APIs from existing API test code.

>>>>

\include{introduction}
\include{case_study/case_study}
\include{spy_rest/spy_rest}
\include{demo_paper/demo_paper}
\include{cisco_study/cisco_study}
\include{controlled_study/controlled_study}
\chapter{Conclusion}
\label{chapter:conclusion}
In this thesis, I focused on improving the state of REST API documentation process. Existing REST API documentation processes rely on largely manual or bespoke approaches that are time consuming and error-prone, specially for evolving REST APIs. Through my research, I aimed to bridge the gap between API source code and its documentation by intercepting usage examples from API test code to automatically generate an always updated API documentation for REST APIs.

I presented a set of five primary research questions related to the topic of REST API documentation. I have addressed the first research question, RQ1, of finding the gap between the current industry practices and existing research on REST API documentation by performing a case study in my first research sub-project. I started by reviewing the existing literature on Web API evolution. From the existing empirical research on Web API evolution, I found that researchers have focused on profiling SOAP based web services to identify the common change patterns. They suggested focusing on the maintenance efforts related to Web API evolution because of its lingering effects. In the existing literature, I found several techniques proposed by researchers related to supporting the evolution of SOAP APIs. Analyzing the existing literature on tool support for Web API evolution, I found tools developed by researchers that primarily rely on custom formatted API specification documents that require a manual effort to generate and maintain. In this sub-project, I have also analyzed the evolution of nine Web APIs in production, selected based on their industry, popularity, maturity, and source code access, to understand the current state of practice around Web API versioning, documentation, and change communication patterns. By comparing and contrasting the findings from the literature review and the industry practices, I observed that it was important to carry out further research to find techniques to support versioning, documentation, and change communication needs for evolving REST APIs.

I addressed RQ2, finding REST API documentation requirements, in my second research sub-project. I started by analyzing the existing literature on API usability to summarize a list of general API documentation requirements as derived by other researchers. Then, I analyzed the existing literature on REST API documentation. Based on these two analyses, I derived the following list of documentation requirements specific for evolving REST APIs: automated, example based, executable, version aware, customizable, reusable, and collaborative. These requirements are used as a background work to address my subsequent research questions.

RQ3, the primary research question of finding a new technique to satisfy the documentation requirements of evolving REST APIs, is addressed in my third and fourth sub-projects. I observed that most existing techniques for REST API documentation relied on custom API specification formats as an input. This approach requires manual effort to generate and maintain the API specification as a separate artifact with the API as the API evolves and fails to satisfy the identified REST API documentation requirements. I proposed a novel technique called SpyREST to automatically intercept and transform example REST API calls into usable API documentation following the identified requirements. An HTTP proxy server is used to collect data from example API calls. The collected data is then synthesized to automatically infer the structural elements of the API such as the version, API resource, HTTP verb, API action, query parameters, request and response body, and HTTP headers. In addition to the structure, executable API examples are automatically re-constructed with realistic data from the recorded API calls. I developed a prototype tool to demonstrate a concrete implementation of the proposed concept. In a preliminary case study with three evolving REST APIs, I have shown that it is possible to use the proposed technique to generate a more accurate documentation of REST APIs compared to their official documentation that are  manually maintained.

I addressed RQ4, an evaluation of the proposed technique from the perspective of REST API developers, in my fourth sub-project using an industrial case study. I presented the proposed technique and the prototype tool to a team of software engineers at Cisco when we were in the process of selecting a documentation method for a suite of new REST APIs to be built. I discussed the selection process behind choosing SpyREST as documentation tool for this project. In this case study, I discussed how the proposed technique of using test code to drive REST API documentation has been leveraged in Cisco to create and maintain an always-updated and accurate documentation for an evolving REST API in production for a period of eighteen months. I gathered feedback from the stakeholders including the API developers, QA engineers, project managers, and customer support engineers to evaluate the merits and shortcomings of the proposed technique based on their production use. The evaluation results show that it is feasible to use API test code to generate and maintain REST API documentation for an evolving API with the identified REST API documentation requirements. From this evaluation, I also found that different stakeholders are able to establish a fast and continuous feedback loop about the API and its documentation while developing the API features when the API documentation is auto updated from executable code. Additionally, from this evaluation, I found a list of limitations of the proposed technique related to the lack of support for using version control features for custom edited API documentation content and automatically generated changelogs showing the differences between API elements of different API versions. Future research can be carried out to address these limitations.

My last research question, RQ5, to evaluate the impact of the proposed technique on REST API client developers, was addressed in my fifth sub-project. I performed a controlled study with 26 professional software engineers with prior REST API experience. The participants were asked to perform tasks using a REST API following its accompanying documentation. I enhanced an existing API documentation with usage examples leveraging the API's test code. This enhancement mimics the outcome of the proposed technique but does not use the proposed tool. Participants were mapped into two groups. One of the groups used the original API documentation and the other group used the enhanced documentation. I provided a web based API explorer so that participants could only focus on the required API inputs instead of writing a full API client application to perform the tasks. I observed that participants using the original API documentation faced similar patterns of obstacles involving the use of correct data types, data formats, HTTP headers, and request body. Participants using the enhanced documentation were able to reduce these obstacles to perform API tasks with a higher success rate, requiring less trial API calls and effort, and a higher satisfaction rating. This provides empirical evidence that the proposed technique has a positive impact on the REST API client developers. I also derived a list of recommendations related to the content of usage examples for REST API documentation that can be used by practitioners as a guideline to prioritize their documentation efforts. From this study, I also observed that participants used more trial API calls for API tasks that involve multiple dependent API calls even while using the enhanced documentation. Future research is needed to generate API documentation with examples for complex API usage scenarios to reduce this obstacle.

@@

\section{Future Work}
Based on my findings from the this research, I identified the following list of potential future research directions.

\begin{itemize}
  \item \textbf{Extension to Support Complex Usage Scenarios.} Future research needs to be carried out to find techniques to automate the process of documenting usage scenarios that require multiple API calls to perform complex API tasks. Future research is also needed to investigate techniques for automatically cross-referencing API elements to allow API client developers with an improved navigation experience while reading the REST API documentation. In Chapter~\ref{chapter:controlled_study}, I discussed that the study participants faced obstacles with performing API tasks that required the use of multiple related API calls to reach a desired outcome. Future research needs to investigate techniques that can be used to provide links to related API elements and usage examples to allow API client developers with an alternative navigation experience.
  \item \textbf{Automated Changelog Generation.} For frequently evolving APIs, future research needs to extend the proposed technique in this thesis to automatically compute human readable changelogs and detect version compatibility to follow a semantic versioning scheme. This would help API client developers with an accurate information about the potential impact of an API update on their own application. In Chapter~\ref{chapter:case_study}, I discussed that REST APIs evolve frequently, often several times per month, and the change profiles contain different types of changes. I also provided a list of change patterns that I observed while studying the change profiles. In Chapter~\ref{chapter:cisco_study}, I discussed that manual effort was required to maintain the API changelog for the studied API. Future research needs to investigate automated changelog generation to help the evolution of REST APIs. The outcome of an automated changelog generation can be applied to automatically infer a semantic version of a REST API based on the inferred compatibility information.
  \item \textbf{Automated API Client Code Generation.} Several tools exist that can convert custom REST API description files to API client code in multiple programming languages. REST API client libraries help developers to leverage various IDE features such as code completion, and suggested data types for API elements. Future research can be carried out to map the SpyREST data model with the custom API description languages to leverage the interception technique to reduce the amount of manual effort required to maintain the custom formatted API descriptions as the API evolves.
  \item \textbf{Extensions of the Interception Based Documentation to non-REST APIs.} The proposed technique can be extended to document usage examples for other stateless APIs outside the realm of REST APIs. The structure of REST APIs are commonly documented using programming language features and formatted comments. Using formatted coments to provide API documentation with executable code examples requires manual effort since no automated feedback is available about the accuracy of the comments. API developers commonly write and maintain API test code that shows the intended way of using the APIs. The API test code provides automated feedback by means of test results. However, API test code may not be a substitute for API documentation. API test code may contain complex setup, tear-down, and verification logic that are not useful for documentation. Also, for closed source applications, organizations may not be able to publish their source code to external API client developers. The concept of intereception as described in Chapter~\ref{chapter:spy_rest} can be applied to other forms of API to generate and maintain API documenation by intercepting REST API tests. For example, several programming languages support concepts such as Aspect Oriented Programming, where custom code can be written to intercept API method calls, that can be used for API documentation. Future research can investigate the feasibility of leveraging the interception technique to source realistic usage examples for local stateless APIs from existing API test code.
